Behind Bars and Beyond: Investigating Socioeconomic Drivers of Incarceration Rates in the United States

### 2024-04-18 paper/paper.qmd
Updated model and model section.
Changed the reference educational category to that of "1-19" since there seemed to be an error with the prior and posterior choices of categories.

- $\mu_i$ is the predicted incarceration rate for a nation given a nation's poverty rate, violent crime rate, unemployment rate, and educational ranking category.
- $\beta_0$ is the coefficient for the intercept.
- $\beta_1$ is the coefficient for the predicted change in the incarceration rate of a nation given a one unit, i.e. one percentage point, increase in the nation's poverty rate.
- $\beta_2$ is the coefficient for the predicted change in the incarceration rate of a nation given a one unit increase in the nation's violent crime rate.
- $\beta_3$ is the coefficient for the predicted change in the incarceration rate of a nation given a one unit, i.e. one percentage point, increase in the nation's unemployment rate.
- $\gamma_i$, for $i = 1, ..., 9$, represents the coefficient for the predicted change in the incarceration rate of a nation given that it is in one of the educational ranking categories, "1-19", ..., "148-175", respectively for each $i$. Notice, this implies that the base intercept assumes that a nation is in the "176-198" educational ranking category.

@@

We can see that, somewhat unexpectedly, increases in the poverty rate negatively correlate with incarceration rates, as do increases in unemployment, assuming all else held constant. On the contrary, as predicted, violent crime rates correlate positively with incarceration rates, all else held constant. As for a nation's educational ranking, it appears that there is no clear relationship between education and incarceration rates as nations in the lowest category of education, "176-198", would see an increase in predicted incarceration rates as they enter categories near the median educational ranking, before then seeing more consistent decreases near the highest ranked nations.

@@

# Calculate the predicted incarceration rate for each nation
predicted_incarceration <- (
  coefficients["(Intercept)"] +                      
  coefficients["poverty_rate"] * predictors$poverty_rate +
  coefficients["violent_crime_rate"] * predictors$violent_crime_rate +
  coefficients["unemployment_rate"] * predictors$unemployment_rate +
  coefficients["education_category1-19"] * as.numeric(predictors$education_category == "1-19") +
  coefficients["education_category20-34"] * as.numeric(predictors$education_category == "20-34") +
  coefficients["education_category35-52"] * as.numeric(predictors$education_category == "35-52") +
  coefficients["education_category53-71"] * as.numeric(predictors$education_category == "53-71") +
  coefficients["education_category72-92"] * as.numeric(predictors$education_category == "72-92") +
  coefficients["education_category93-110"] * as.numeric(predictors$education_category == "93-110") +
  coefficients["education_category111-125"] * as.numeric(predictors$education_category == "111-125") +
  coefficients["education_category126-147"] * as.numeric(predictors$education_category == "126-147") +
  coefficients["education_category148-175"] * as.numeric(predictors$education_category == "148-175")
)

@@

```{r}
#| label: fig-rhat
#| echo: false
#| layout-ncol: 2
#| fig-pos: H

>>>>

- $\mu_i$ is the predicted incarceration rate for a nation given a nation's poverty rate, violent crime rate, unemployment rate, and educational ranking category.
- $\beta_0$ is the coefficient for the intercept.
- $\beta_1$ is the coefficient for the predicted change in the incarceration rate of a nation given a one unit, i.e. one percentage point, increase in the nation's poverty rate.
- $\beta_2$ is the coefficient for the predicted change in the incarceration rate of a nation given a one unit increase in the nation's violent crime rate.
- $\beta_3$ is the coefficient for the predicted change in the incarceration rate of a nation given a one unit, i.e. one percentage point, increase in the nation's unemployment rate.
- $\gamma_i$, for $i = 1, ..., 9$, represents the coefficient for the predicted change in the incarceration rate of a nation given that it is in one of the educational ranking categories, "20-34", ..., "176-198", respectively for each $i$. Notice, this implies that the base intercept assumes that a nation is in the "1-19" educational ranking category.

@@

We can see that, somewhat unexpectedly, increases in the poverty rate negatively correlate with incarceration rates, as do increases in unemployment, assuming all else held constant. On the contrary, as predicted, violent crime rates correlate positively with incarceration rates, all else held constant. As for a nation's educational ranking, it appears that there is no clear relationship between education and incarceration rates as nations in the highest category of education, "1-19", would see a much larger increase in predicted incarceration rates as they enter categories near the median educational ranking, before then seeing smaller increase near the lowest ranked nations.

@@

# Calculate the predicted incarceration rate for each nation
predicted_incarceration <- (
  coefficients["(Intercept)"] +                      
  coefficients["poverty_rate"] * predictors$poverty_rate +
  coefficients["violent_crime_rate"] * predictors$violent_crime_rate +
  coefficients["unemployment_rate"] * predictors$unemployment_rate +
  coefficients["education_category20-34"] * as.numeric(predictors$education_category == "20-34") +
  coefficients["education_category35-52"] * as.numeric(predictors$education_category == "35-52") +
  coefficients["education_category53-71"] * as.numeric(predictors$education_category == "53-71") +
  coefficients["education_category72-92"] * as.numeric(predictors$education_category == "72-92") +
  coefficients["education_category93-110"] * as.numeric(predictors$education_category == "93-110") +
  coefficients["education_category111-125"] * as.numeric(predictors$education_category == "111-125") +
  coefficients["education_category126-147"] * as.numeric(predictors$education_category == "126-147") +
  coefficients["education_category148-175"] * as.numeric(predictors$education_category == "148-175") +
  coefficients["education_category176-198"] * as.numeric(predictors$education_category == "176-198") 
)

@@

```{r}
#| label: fig-rhat
#| echo: false
#| fig-pos: H
